Indirect Conflicts: An Exploration and Discussion of Tools, Process, and Developer Insight

### 2013-10-03 iced-2013/paper.tex
ICED 2nd draft almost done. Needs one read.

\usepackage{balance}
\usepackage{float}

@@

\section{Results}
\label{sec:results}

We now present our results of both the interviews and surveys conducted in regards to our research questions
outlined in Section~\ref{sec:intro}. We first present and discuss the general nature of indirect conflicts in
terms of when, what types, and how often indirect conflicts occur. We next show the findings for the process
related research questions in regards to how developers react to, processes to prevent, and reactionary steps
to indirect conflicts.

\subsection{The Nature of Indirect Conflicts}

Although a detailed manual inspection of a project's lifetime and the indirect conflicts it experiences throughout
its development is beyond the scope of this paper, we first wanted to explore the general nature of indirect 
conflicts. This being the case we explored, through our interviews and surveys, what software developers believed
to be the case when it came to how indirect conflicts occur, when and how often they occur, and the types of 
software object's that incur indirect conflicts.

\begin{description}
	\item[~\ref{itm:rq1}] \textit{How do indirect conflicts occur?}
\end{description}

From the interviewed participants, 63\% of developers believe that a large contributing factor to the cause
of indirect conflicts comes from the changing of a software object's contract. Object contracts are, in a sense,
what a software object guarantees, meaning how the input, output, or how any aspect of the object is garunteed
to work; made famous by Eiffle Software's~\footnote{http://www.eiffel.com/} ``Design by Contract''\texttrademark. 
In light of object contracts, 73\% of interviewed developers gave examples of indirect conflicts they had experienced
which stemmed from not understanding the far reaching ramifications of a change being made to an object contract
towards the rest of the project. Of those 73\%, examples were broken down as follows. 21\% of examples dealt
with the changing of legacy code, with one developer saying ``legacy code does not change because developers 
are afraid of the long range issues that may arise from that change''. 36\% of interviewed developers mentioned 
that indirect conflicts occur when an update to an external project, library, or service occurs and the resulting 
integration fails noting that major releases of libraries can be a large issue with one developer saying 
``their build never breaks, but it breaks ours''. Finally, 29\% of developers interviewed explained that changes to
databases are often a large source of indirect conflicts saying ``[schema changes] break stuff all over the place''.

\begin{description}
	\item[~\ref{itm:rq2}] \textit{How often do indirect conflicts occur and at what times in a project do they occur more frequently?}
\end{description}

From the interviewed participants, 58\% of developers explained that indirect conflicts occur all the time in
their development life cycle with a minimum occurrence of once a week. Of those 58\%, developers noted that a
lot of their development time is put into understanding code to avoid indirect conflicts, more serious issue tend
to happen once a month, and that the conflicts that do occur tend to be quite unique from each other.

@@

For those developers interviewed, we asked how often indirect conflicts occur in their software development experience,
how often they occur before and after a project's first release, as well as how often indirect conflicts occur late
in a project's lifetime. Table~\ref{tab:often} shows the results of these questions.

@@

Of those surveyed that could not give direct answers to the questions in Table~\ref{tab:often}, the more common answers were 
as follows: ``People rushing to meet [the release date] is where most of my indirect conflicts occur.'', ``Indirect conflicts
after a release depends on how well the project was built at first'', ``They tend to slow down a bit after a major release, 
unless the next release is a major rework.'', and ``Conflicts continue to occur at roughly the same rate after the first release,
with spikes during large revamps or the implementation of cross-cutting new features.''.

\begin{description}
	\item[~\ref{itm:rq3}] \textit{What software artifacts and environments are susceptible to indirect conflicts?}
\end{description}

From the developers interviews, 9 were currently working with large scale database applications and all 9 listed database
schema as a large source of indirect conflicts. Out of the 5 developers interviewed that currently work on either software
library projects or in test, all 5 said that methods or functions were the root of their indirect conflict issues. Some
other notable artifacts were user interface displays for web development, and full components in component base game
architecture.

@@

\subsection{Processes of Indirect Conflicts}

The results that follow pertain to software development process and its relation to indirect conflicts. We explore
how developers react to indirect conflicts, which stakeholders are responsible for their fixes, and if any development
processes can be used to prevent or catch indirect conflicts.

\begin{description}
	\item[~\ref{itm:rq4}] \textit{How do developer prevent indirect conflicts?}
\end{description}

Of the developers interviewed, 3 main points of interest were found in preventing indirect conflicts. The first was
the use of ``Design by Contract''\texttrademark, meaning that software objects have certain guarantees about their behavior
in terms of input, output, and internal workings. This methodology can be used as a type of documentation towards awareness of
a software object. One developer stated that ``design by contract was invented to solve this problem and it does it quite well'', 
while another noted that software  object contracts do solve the problem in theory, but that doesn't mean that problems don't 
still occur in practice. The second point was the add and deprecate methodology. 21\% if developers interviewed mentioned 
that the development process changes once a stable point of a project is met and the ``add and deprecate model [prevents] a lot 
of these issues''. Add and deprecate meaning instead of editing code, the developer simply clones code (if needed) and edits the clone
while slowly phasing out the out code in subsequent releases or other development cycle needs. This allows a stoppage of indirect conflicts
by allowing developers to use the older versions of software objects which remain unchanged. Lastly, pure developer experience was mentioned
as a source to prevent indirect conflicts. 37\% of developers mentioned that when planning code changes, either a very experienced member
of the project was involved in the planning and has duties to foresee any indirect conflicts that may arise, or that while implementing
a code change developers must use their personal knowledge to predict where indirect conflicts will occur.

Of the developers surveyed who could give positive identification to preventative measures taken to avoid indirect conflicts (37 developers
total), 27\% said that individual knowledge of the code base and their impact of code change was used, 59\% mentioned some form of ``Design
by Contract'' or the testing of a methods contract (which could be viewed as catching indirect conflicts), and 14\% said that add and
deprecate was used in their projects to avoid indirect conflicts.

\begin{description}
	\item[~\ref{itm:rq5}] \textit{How do developers catch indirect conflicts?}
\end{description}

Of the 19 developers interviewed, 69\% mentioned forms of testing (unit, integration, etc) as the major component of catching indirect
conflict issues, subscribing to the idea of ``run the regression and integration tests and just see if anything breaks''. 31\% of developers 
said build processes (either nightly builds or building the project themselves), and others mentioned code reviews
while those dealing with a user interface mentioned user complaints from run time testing.

Of the 78 developers survey, 49\% mentioned forms of testing as the major tool used to catch indirect conflicts, 33\% said build processes,
while 31\% used work their IDE or IDE plug-ins to catch indirect conflicts. Other developers surveyed also mention code review process (14\%)
and project expertise (6\%) as factors of catching indirect conflicts.

\begin{description}
	\item[~\ref{itm:rq6}] \textit{How to developers react to and solve indirect conflicts?}
\end{description}

@@

Of the developers surveyed, 37\% confirmed that they used some type of source control management tools to help them solve their indirect
conflicts while 23\% said they used native IDE tools and 21\% said they use features of the language's compiler and debugger in order
to solve indirect conflicts. Interestingly, only 13\% of developers mentioned a form of communication with other developers in aid to
solving these conflicts and only 4\% mentioned the reading of formal documentation.

\subsection{Tools for indirect Conflicts}

\begin{description}
	\item[~\ref{itm:rq7}] \textit{When are preemptive indirect conflict tools best used and what would developers want from them?}
\end{description}

The developers who were interviewed had a major concern with any preemptive tool in that the amount of false positives
provided by the tool may render the tool useless. Developers said ``this could relate to information overload'', ``this would
be a real challenge with the number of dependencies'', ``I only want to know if it will break me'', and ``it depends on how
good the results are in regards to false positives''. These concerns combined with RQ2 suggest that preemptive tools are
best used in unstable phases on the project when indirect conflicts are more likely to happen, thus reducing the number of false
positives. One developer in particular noted ``You will spend all your time getting notified and reviewing instead of implementing''.
Interviewed developers also suggested that their proper software development process are already in place to
catch potential issues before they arise such as code review, individual knowledge, IDE call hierarchies (static language analysis),
or communication within the project. Finally, most interviewed developers mentioned that their current work flow does
not subscribe to the ideas of preemptive prevention and that they just ``work until something breaks''.

@@

\begin{description}
	\item[~\ref{itm:rq8}] \textit{What kinds of tools do developers want to help them handle and solve indirect conflicts?}
\end{description}

The developers who were surveyed, the following list represents the more common suggestions for how an indirect conflict
resolution tool should be implemented.

@@

\section{Discussion}
\label{sec:disc}

We discuss the results of Section~\ref{sec:results} as follows: first we look at and discuss the nature of indirect conflicts by addressing~\ref{itm:rq1} 
-~\ref{itm:rq3}, we proceed to talk about how developers structure their work flow around indirect conflicts currently by
examining the results of~\ref{itm:rq4} -~\ref{itm:rq6}, and we lastly investigate why developers have not had much interest in tools built previously
for indirect conflicts and what we can do moving forward to accommodate developer work flow with potential tools by looking at~\ref{itm:rq7} 
and~\ref{itm:rq8}.

\subsection{The Nature of Indirect Conflicts}

In order to build solution tools for indirect conflicts, we must first understand their nature; what causes them and how often do they 
occur. From the results we see in~\ref{itm:rq1}, a common theme of awareness is obvious. Developers cause indirect conflicts because they do not
fully understand the ramifications of their source code change. This was especially apparent when developers were dealing with legacy code.
Developers identified the changing of older code to be quite prone to indirect conflicts as they were not sure where the code was being used
anymore. We would think this can be a simple fix with static analysis tools of source
code to allow developers to identify which parts of the project they are affecting with a change. However, as we found out this is not always
the easy fix. For starters, not all programming languages are subject to strong static analysis if any, while some projects involve many
cross language dependencies which also breaks down any potential for static analysis on these software dependencies. 
Instances of upgrading libraries to newer versions was a common cause of indirect conflicts, and since these may be binary files,
static analysis of the changing code is not always possible. The other side of upgrading libraries also posed a problem. Developers who were often
tasked with updating older libraries had no knowledge of what other outside projects used these libraries and were therefore unable to understand
how they may affect these outside projects with their changes.

We see from~\ref{itm:rq3} that developers have shown that the human factor involved with indirect conflicts cannot be ignored.
From Table~\ref{tab:env}, is can clearly be seen that as a project becomes larger and more and more developers are added, even to the point
that multiple teams have to be formed, indirect conflicts become more likely to occur. However, this is not to say that indirect conflicts
do not occur at the lower number of developer level. Even when developing individual projects, 43\% of developers agreed or strongly agreed
that indirect conflicts are likely to occur. Once more than one developer is associated with a project, or a project has multiple teams, 
67\% and 74\% of developers agreed or strongly agreed that indirect conflicts are likely to occur. So even when developers are working alone,
they do not have enough awareness of their own code as to how a software change may affect another portion of their project. This can be due to
any of the reasons mentioned above. The problem also seems to be compounded once multiple developers and teams are involved in parallel development
schemes. The lack of awareness as to other developers work in a project again comes to light. This lacking is increased as project size trends up.
It would appear possible that the addition of software developers to a project, which seems to compound the likely hood of indirect
conflicts, developers would be able to leverage other developer expertise of the project to help avoid or solve indirect conflicts.
However, we will see in Section~\ref{sec:pd} that most developers do not leverage their social connections within a project to help
solve indirect conflicts.

From ~\ref{itm:rq2}, we can make the deduction that indirect conflicts are more likely to occur early in a projects development life cycle as opposed
to later. We used to notion of the first major release to make the distinction between early and late in a project but the same hypothesis
could be applied to iterative development where during the earlier stages of the current phase could be considered more probable for indirect
conflict to occur. In Table~\ref{tab:often}, we see that as the project becomes more mature, a drop off in the frequency of indirect
conflicts occurs. Developers surveyed answered that daily and weekly indirect conflict frequencies trail off past the project's first release and 
those developers interviewed mentioned a point in the project, the ``mature point'', in which indirect conflicts become less frequent after.
This would suggest that, as will be seen in Section~\ref{sec:td}, preemptive indirect conflict tools would be more useful before this point as the
number of false positives will have a higher chance of being true positives. This ``mature point'' may be cause for one of two reasons. The
first would be that the project's source code has stabilized quite a bit in terms of large changes not occurring as much. This could be a result
of a project being released to a customer or because it is now being used as part of many other projects. The second would be that a change in
development process has occurred in that source code is not being changed so much as newer code is being added or old code is being deleted. This
would cause a slow in indirect conflicts as developers interviewed said that changes to code, not the addition of or deletion of code is prone
to indirect conflicts. This two possibilities are not mutually exclusive. In the end, developers survey pointed out that before the ``mature
point'' indirect conflicts are like to occur daily and weekly (32\% and 18\%) as opposed to after the ``mature point'' when developers
said they are more likely to occur on a monthly or even yearly frequency (18\% and 12\%).

>>>>

%\usepackage{balance}
\usepackage{float}

@@

\section{Results and Discussion}
\label{sec:results}

We now present our results of both the interviews and surveys conducted in regards to our 3 research questions
outlined in Section~\ref{sec:intro}. We list each finding followed by our qualitative and quantitative results
from which we draw the finding which are then followed by discussion.

\subsection{What is the nature of indirect conflicts?}

Although a detailed manual inspection of a project's lifetime and the indirect conflicts it experiences throughout
its development is beyond the scope of this paper, we first wanted to explore the general nature of indirect 
conflicts. This being the case, we explored, through our interviews and surveys, what software developers believed
towards to how indirect conflicts occur, when and how often they occur, and the types of 
software object's that incur indirect conflicts.

\begin{description}
	\item[F1\namedlabel{itm:f1}{F1}] Indirect conflicts are caused by a lack of understanding and awareness of source code changes which are brought on
						by the manipulation of software contracts. This lack of awareness and understanding is compounded by having a multi developer
						or multi team environment.
\end{description}

From the interviewed participants, 63\% of developers believe that a large contributing factor to the cause
of indirect conflicts comes from the changing of a software object's contract. Object contracts are, in a sense,
what a software object guarantees, meaning how the input, output, or how any aspect of the object is guaranteed
to work; made famous by Eiffle Software's~\footnote{http://www.eiffel.com/} ``Design by Contract''\texttrademark. 
In light of object contracts, 73\% of interviewed developers gave examples of indirect conflicts they had experienced
which stemmed from not understanding the far reaching ramifications of a change being made to an object contract
towards the rest of the project. Of those 73\%, examples were broken down as follows. 21\% of examples dealt
with the changing of legacy code, with one developer saying ``legacy code does not change because developers 
are afraid of the long range issues that may arise from that change''.

From these listed results, we see a common theme of awareness is obvious. Developers cause indirect conflicts because they do not
fully understand the ramifications of their source code change, especially when a software object's contract is being changed. 
This was largely apparent when developers were dealing with legacy code,
or another developer's code in general. The non-understanding of how implementation decisions were made is a large factor and that changing
older code was quite prone to indirect conflicts as they were unsure of where the code was being used. This presents an immediate problem regarding
indirect conflicts in software maintenance and we would think the simple fix would be static analysis tools for source
code to allow developers to identify which parts of the project they are affecting with a change. However, as we found out this is not always
the easy fix. For starters, not all programming languages are subject to strong static analysis if any, while some projects involve many
cross language dependencies which also breaks down any potential for static analysis on these software dependencies. Another potential solution
to this issue could be formal documentation such as UML. However, from Petre~\cite{Petre:2013:UP} we see that UML is not used so much in practice.
These short comings leave the door open for other type of pro-developer documentation or tools to allow quick understanding of code changes and
how they may impact cross team, cross project, or other dependencies. At any rate, it is clear that developers do not posses the tools needed
to quickly evaluate a software change (especially when contracts change) from a wide spectrum of dependencies that may not have the ability to
be statically analyzed.

\begin{description}
	\item[F2\namedlabel{itm:f2}{F2}] Indirect conflicts, while occurring quite frequently throughout a development cycle,
						occur more often before a project reaches its ``mature point'' rather than after for each development cycle.
\end{description}

From the interviewed participants, 58\% of developers explained that indirect conflicts occur ``all the time'' in
their development life cycle with a minimum occurrence of once a week. Of those 58\%, developers noted that a
lot of their development time is put into understanding code to avoid indirect conflicts, with more serious issues tending
to occur once a month, and that the conflicts that do occur tend to be quite unique from each other. Right away we can see a serious issue with the standard
development work flow. Developers are not actually spending their time developing, rather they are spending large amounts of time trying to avoid indirect
conflicts. This shows that indirect conflicts are a serious threat to developer productivity and are a problem worth study and investigation.

@@

For those developers surveyed, we used this notion of a mature or stable point to ask how often indirect conflicts occur in their software development experience,
how often they occur before and after a project's first release (mature point), as well as how often indirect conflicts occur late
in a project's lifetime. Table~\ref{tab:often} shows the results of these questions.

@@

From the interviews and the surveys combined we can make the deduction that indirect conflicts are more likely to occur early 
in a projects development cycle as opposed to later. This must be a major concern for any indirect conflict tools moving forward. Since many
of the tools listen in Section~\ref{sec:intro} implement some sort of filtering system or taxonomy to avoid information overload, this
finding should be included in future workings. Knowing that indirect conflicts occur more frequently early in a development cycle will let
these tools know that those reported later in the development life cycle have a high probability (compared to without this knowledge) of
being a true positive. This is a major concern when dealing with information overload and to provide developers with what they may find
relevant to their work. The results in in Table~\ref{tab:often} also suggest that preemptive indirect conflict tools may be more accurate
towards the mature point of a project, but have potentially more use earlier where the indirect conflicts are more likely to occur.
While what the ``mature point'' of a project is could be speculated upon, we suggest it is caused for one of two reasons, The
first would be that the project's source code has stabilized quite a bit in terms of large changes not occurring as much. This could be a result
of a project being released to a customer or because it is now being used as part of many other projects. The second would be that a change in
development process has occurred in that source code is not being changed so much as newer code is being added or old code is being deleted. This
would cause a slow in indirect conflicts as developers interviewed said that changes to code, not the addition of or deletion of code is prone
to indirect conflicts. In the end, developers survey pointed out that before the ``mature
point'' indirect conflicts are like to occur daily and weekly (32\% and 18\%) as opposed to after the ``mature point'' when developers
said they are more likely to occur on a monthly or even yearly frequency (18\% and 12\%).

Of those surveyed that could not give direct answers to the questions in Table~\ref{tab:often}, the more common answers were 
as follows: ``People rushing to meet [the release date] is where most of my indirect conflicts occur.'', ``Indirect conflicts
after a release depends on how well the project was built at first'', ``They tend to slow down a bit after a major release, 
unless the next release is a major rework.'', and ``Conflicts continue to occur at roughly the same rate after the first release,
with spikes during large revamps or the implementation of cross-cutting new features.''. These results can be argued to show that the occurance
of indirect conflicts can largely depend on the management, and deadlines of projects. These are important factors to consider when trying
to define a ``mature point''. A large refactoring or rewrite can reset the ``mature point'' of a project and should be considered for
any taxonomies created for preemptive indirect conflict tools.

@@

From Table~\ref{tab:env}, is can clearly be seen that as a project becomes larger and more and more developers are added, even to the point
that multiple teams have to be formed, indirect conflicts become more likely to occur. However, this is not to say that indirect conflicts
do not occur at the lower number of developer level. Even when developing individual projects, 43\% of developers agreed or strongly agreed
that indirect conflicts are likely to occur. Once more than one developer is associated with a project, or a project has multiple teams, 
67\% and 74\% of developers agreed or strongly agreed that indirect conflicts are likely to occur. These results show that indirect conflicts
seem to be compounded by a multiple human factor found in parallel development. The lack of awareness as to other developers work in a project 
again comes to light. We would think that developers would leverage the other developer's knowledge in their projects to avoid indirect conflicts
but as we will see most developers tend not leverage their social connections within a project to help solve indirect conflicts. This could be
a potential issue of social structure in that developers simply do not have the means or will to talk to their colleagues, or it could be that
developers heavily leverage the notion of software stigmergy as shown by Bolici et. al~\cite{Bolici}. Creating effective communication around
technical dependencies between developers and teams through either explicit or implicit means appears to be a large task to accomplish but a
carsickness task as shown by these result being that static analysis breaks down at these levels.

\begin{description}
	\item[F3\namedlabel{itm:f3}{F3}] Indirect conflicts are largely brought on by the changing of interfaces to software structures. The most common type of structures
						to cause indirect conflicts are methods/functions and database schemas.
\end{description}

From the developers interviewed, 9 were currently working with large scale database applications and all 9 listed database
schema as a large source of indirect conflicts with developers saying ``[schema changes] break stuff all over the place''.
Out of the 5 developers interviewed that currently work on either software
library projects or in test, all 5 said that methods or functions were the root of their indirect conflict issues. 
36\% of interviewed developers mentioned that indirect conflicts occur when an update to an external project, library, or service occurs and the resulting 
integration fails noting that major releases of libraries can be a large issue with one developer saying 
``their build never breaks, but it breaks ours''. Some
other notable artifacts were user interface displays for web development, and full components in component base game
architecture.

From these results, we can see the evidence that the problemed areas of software architecture in regards to indirect conflicts are
the interfaces between structures, mainly being methods and database schemas. This is good news for preexisting indirect conflict tools as
many of them have focused at the signature level of methods as a spot for technical dependency changes. However, there has been little focus on indirect
conflicts at the relation database schema level~\cite{Maule:2008:IAD}, possibly due to the inherit difficult nature
of analysis of a database schema in relation to source code structure. Developers have given researchers an area of indirect conflicts that have been
over looked and we should be making an effort to provide answers in this area. Also, note here that developers while in F1 showed that changes to software
object contracts were the concern, the focus here was shifted to the interface. While the interface does make up a large component of a software object contract,
it does not fully encompass it. Precondition and postcondition checks, which are part of the object contract are performed in the body of a method and are largley overlooked
by indirect conflict tools. This may be a source of many false negatives from these tools.

Instances of upgrading libraries to newer versions was a common cause of indirect conflicts, and since these may be binary files,
static analysis of the changing code is not always possible. Developers who were also
tasked with updating older libraries had no knowledge of what other outside projects used these libraries and were therefore unable to understand
how they may affect these outside projects with their changes. However, it appears that when interfaces of methods change through software
evolution, indirect conflicts are more prone to occur. This result again shows that even at the library or binary level, any type of support tools for indirect
conflicts should focus at the interface level of change.

\subsection{What types of process and tools are being used by developers in regards to indirect conflicts?}

The findings that follow pertain to software development process and its relation to indirect conflicts. We explore
how developers react to indirect conflicts, which stakeholders are responsible for their fixes, and if any development
processes can be used to prevent or catch indirect conflicts.

\begin{description}
	\item[F4\namedlabel{itm:f4}{F4}] Developers use a combination of object contract understanding, the add and deprecate model, as well as expert 
																	 knowledge in order to prevent indirect conflicts.
\end{description}

Of the developers interviewed, 3 main points of interest were found in preventing indirect conflicts. The first was
the use of ``Design by Contract''\texttrademark, meaning that software objects have certain guarantees about their behavior
in terms of input, output, and internal workings. This methodology can be used as a type of documentation towards awareness of
a software object. One developer stated that ``design by contract was invented to solve this problem and it does it quite well'', 
while another noted that software object contracts do solve the problem in theory, but that doesn't mean that problems don't 
still occur in practice.  Since ``Design by Contract''\texttrademark ensures that software artifacts, while they can be changed, 
should always behave in a way that is understood by other developers, languages should just directly support ``Design by Contract''\texttrademark,
however, this is not the case. One developer said that ``is it too much of a burden'' in regards to strictly making the contract
apart of the source code. But why is it not such a burden is write separate unit tests that are normally designed around testing
the contract? There is obviously a trade off between having rules built into the source versus having external tests to check if rules
were implemented correctly. The question should become which is more effective for software development in terms of bugs produced 
from contract breaks or bugs produced from convoluted code. This paper cannot begin to attempt to answer this question.

The second point was the add and deprecate methodology. 21\% if developers interviewed mentioned 
that the development process changes once a stable point of a project is met and the ``add and deprecate model [prevents] a lot 
of these issues''. Add and deprecate meaning instead of editing code, the developer simply clones code (if needed) and edits the clone
while slowly phasing out the out code in subsequent releases or other development cycle needs. This allows a stoppage of indirect conflicts
by allowing developers to use the older versions of software objects which remain unchanged. From the interviews, it seems this is a largely
effective way of preventing indirect conflicts, however, it was mostly used for libraries and methods only. In some cases such as development
in a large database project, the add and deprecate model simply does not work as all code must run off the newest changes. So while the
add and deprecate model is certainly a good fix for heavily version-ed software, especially when given to 3rd party developers, it does
not necessarily work on internal projects which need to cooperate on which versions of methods or schemas they use.

Lastly, pure developer experience was mentioned
as a source to prevent indirect conflicts. 37\% of developers mentioned that when planning code changes, either a very experienced member
of the project was involved in the planning and has duties to foresee any indirect conflicts that may arise, or that while implementing
a code change developers must use their personal knowledge to predict where indirect conflicts will occur. While most interviewed developers
had such an expert in their software projects, they admitted to not being able to foresee all indirect conflicts that may occur, especially
when talking about external teams or projects who may or may not use their final product. This raises the question of how much planning
by expert developers should really go into avoiding indirect conflicts? Dromey~\cite{} raises the debate of prevention versus cure of software
defects and how difficult a problem it is to measure. Should developers be using certain tools to plan their changes or should they be even
planning at all? Which method provides the best software quality for the minimum effort. These are not easy questions to answer but may
be key in improving developer productivity when concerning indirect conflicts.

To confirm our findings, the developers surveyed who could give positive identification to preventative measures taken to avoid indirect conflicts (37 developers
total), 27\% said that individual knowledge of the code base and their impact of code change was used, 59\% mentioned some form of ``Design
by Contract'' or the testing of a methods contract (which could be viewed as catching indirect conflicts), and 14\% said that add and
deprecate was used in their projects to avoid indirect conflicts.

Even with proper process and expert knowledge, it is still improbable to predict every indirect conflict, and thus, conflicts 
continue to occur. However, developers did not seem overly concerned that these three items do not totally prevent indirect conflicts. This
is because most developers ascribe to the idea of ``I work until something breaks'', or taking the curative approach. This shows that developers are somewhat content with
the level of prevention that happens for indirect conflicts, however, this could be an issue of developers simply following old standards.
It seems that the cure approach to indirect conflicts is heavily used in industry, but is it better than prevention?
The lack of preventative process being used is probably due to the difficult nature of predicting indirect conflicts. These results have mostly
shown that developers would rather spend their time actually writing the software than trying to prevent only potential issues. Time is
more effectively spent, in the developer's eyes, developing and fixing problems as needed than putting excess time into the prevention of potential errors.
This is an important consideration for moving forward with indirect conflict tools and research.

\begin{description}
	\item[F5\namedlabel{itm:f5}{F5}] Developers use unit and integration testing to examine use case coverage in order to detect indirect conflicts.
\end{description}

Of the 19 developers interviewed, 69\% mentioned forms of testing (unit, integration, etc) as the major component of catching indirect
conflict issues, subscribing to the idea of ``run the regression and integration tests and just see if anything breaks''. 31\% of developers 
said build processes (either nightly builds or building the project themselves), and others mentioned code reviews
while those dealing with a user interface mentioned user complaints from run time testing. It appears that most indirect conflicts are currently
detected by using what some would call proper software development practices. The words ``use case coverage'' were constantly being used by developers
when expressing how proper unit and integration tests should be written. Developers expressed that with proper use case coverage, most if
not all indirect conflicts should be caught. This leads us to believe that as long as the contract of a software structure is understood,
proper tests should be able to be written which will catch most if not all indirect conflicts. This is another example of developers subscribing
to the work until something breaks philosophy in that if proper tests are written, the errors will be caught after they break something.
However, problems occur when a software contract is
not properly understood and use cases are not tested for or examined. Once an object is not properly
tested, the negative consequences may not be caught until later in a project with the potential for more damage to be caused. This is the major
negative result of work until something breaks. The break detectors need to be properly written. Some languages
such as Eiffle and the Java Modeling Language~\footnote{http://www.eecs.ucf.edu/~leavens/JML//index.shtml}
enforce object contracts to be well understood through the ideas of preconditions and postconditions of structures which is an example of preemptive
conflict detection. However, the trade off between external tests and in-line contracts is still in place as previously mentioned.

Of the 78 developers survey, 49\% mentioned forms of testing as the major tool used to catch indirect conflicts, 33\% said build processes,
while 31\% used work their IDE or IDE plug-ins to catch indirect conflicts. Other developers surveyed also mention code review process (14\%)
and project expertise (6\%) as factors of catching indirect conflicts. 
Developers said that the build process is simply there to ensure a, at least, semi-functioning product. They use
the build process more to catch compile time errors witch prevent the product from being used. These types of build breaking indirect
conflicts do occur but for the most part developers posses tools capable to build their solutions in order to catch these problems before
the code change is added to the project. Therefore, unit and integration testing appears to be the largest factor in catching indirect conflicts
as long as they are properly written and contracts are understood.

\begin{description}
	\item[F6\namedlabel{itm:f6}{F6}] Developers prefer to spend time debugging real problems opposed to preventing potential problems.
\end{description}

@@

Here, the generic approach
brought forth by developers was seen to be history checking. Developers would look into source code change logs and build or test failure
logs in order to see what is broken and what might have changed recently that could be affecting the break. This gives the impression that
developers already have the tools to solve indirect conflicts. However, this leads to the question: how quickly and
efficiently are they able to solve them and is the time spent debugging outweighing the potential time spent preventing such errors?
While each developer gave some examples on solving indirect conflicts and the times needed, they all agreed that to be able to solve them
quicker would be a huge benefit. This again shows the developer preference towards work until something breaks as opposed to prevent the
error in the first place. While the debate of prevention versus cure goes on, possible solutions as to solving
indirect conflicts quicker are the automatic analysis of source code
change logs to detect what changes will affect a certain part of a project's code. Luckily, some of this solution has already been implemented
by several researchers in both delta debugging~\cite{Zeller:2002:ICC} and program slicing~\cite{Weiser:1982:PUS}. The fact that researchers
have had such success in the cure or debugging side of indirect conflicts could be due to the lack of indirect conflict. Once a problem has
occurred, there is no chance of it being a false positive. Academics do not have to create taxonomies to filter results and developers do
not need to spend time working on only potential issues. Their time is more effectively, in their eyes, spent developing than speculating.

Of the developers surveyed, 23\% also said they used native IDE tools and 21\% said they use features of the language's compiler and debugger in order
to solve indirect conflicts. Interestingly, only 13\% of developers mentioned a form of communication with other developers in aid to
solving these conflicts and only 4\% mentioned the reading of formal documentation. This shows is the serious lack of static analysis tools
available for developers in debugging as many projects and languages simply do not support it (adding cause for non static analysis tools moving
forward), but more importantly, that developers tend to not communicate or use formal documentation for debugging indirect conflicts. Petre~\cite{Petre:2013:UP},
as already mentioned, showed that it is not industry practice to use UML, which could open to floor to more informal means such as online forums. The none
developer communication, as again already mentioned, could be addressed by either more formal communication structures or by developer stigmergy 
in their projects if needed.

\subsection{How should indirect conflict tools work to best benefit developers?}
Ultimately, the goal of this research would be to produce any sort of tool that could aid developers in any stage (prevention, detection, or
solving) of indirect conflicts. In order to address what types of tools would be best, we analyze the results from our previous research
questions along with the proceeding results.

\begin{description}
	\item[F7\namedlabel{itm:f7}{F7}] It is extremely difficult, if not impossible, to create a preemptive indirect 
																	 conflict tool which can detect indirect conflicts with a low rate of false positives. Preemptive tools 
																	 should be exploratory in nature.
\end{description}

The developers who were interviewed had a major concern with any preemptive tool in that the amount of false positives
provided by the tool may render the tool useless. Developers said ``this could relate to information overload'', ``this would
be a real challenge with the number of dependencies'', ``I only want to know if it will break me'', and ``it depends on how
good the results are in regards to false positives''. These concerns combined with RQ1 suggest that preemptive tools are
best used in stable phases on the project when indirect conflicts are less likely to happen, thus reducing the number of false
positives reported. One developer in particular noted ``You will spend all your time getting notified and reviewing instead of implementing''.
Interviewed developers also suggested that their proper software development process are already in place to
catch potential issues before they arise such as code review, individual knowledge, IDE call hierarchies (static language analysis),
or communication within the project. All of this again shows that developers would prefer to subscribe to the 
``work until something breaks'' philosophy. Again, it is up for debate whether or not this is the correct approach the developers should
be taking (effectiveness of prevention versus cure), but it is the one they currently prefer.

@@

Developers were very quick to identify this problem which has
been a common occurrence among any preemptive indirect conflict tool built. The main issue with false positives is the detection of changes
which frequently cause indirect conflicts. Change types are very possible to detect through software static analysis~\cite{Fluri:2007:CDT} (when possible), 
but even those that are
prone to causing indirect conflicts do not cause them all the time. This is quite evident in Table~\ref{tab:pre} where developers identified
that there is an even distribution of software changes that could cause indirect conflicts and as to if they would want to be made aware of
those changes. The only two exceptions which stand out are that of method signature changes and main algorithm changes. The problem here however,
is that method signature changes will be caught by IDE errors, nightly builds, unit testing, integration testing and so on, while main algorithm
changes are very difficult to detect. This all leads to the hypothesis that preemptive indirect conflict tools cannot be built without a very
large number of false positives being reported at any given time. Even if developers, as per Table~\ref{tab:pre}, want to be alerted to 
changed preconditions in methods most times if not always, there is still a large chance that that changed precondition will actually not
introduce an indirect conflict in the source code. This leads to a large number of false positive being reported, which leads to information
overload, and eventually to developers abandoning the tool.

%In order to prevent or catch indirect conflicts with an automated tool without introducing a large amount of false positives, would be to
%have an automatic understanding of how the project's code is suppose to work. From the halting problem~\cite{Burkholder:1987}, 
%we know it is an impossible task to write one program that can verify the correctness of another. This leaves us with the only option of some kind of
%human intervention in the matter which leads us back to use case coverage. As was stated before, if the developer has a complete understand
%of the use cases of a particular code artifact, then they can write proper tests to ensure those use cases are always met and not broken which
%would be the best form of preventing and catching indirect conflicts.

Given the above challenges, some indirect conflict tools have had some general success. The key difference with these tools is that they
contain the idea of developer exploration. The idea of only presenting information to users on request is the complete opposite of the preemptive
notion of notifying developers to potential upcoming problems. Here, the idea that a user can provide some sort of
input, a sort of context, is a powerful tool for reducing false positives. We have really just moved the false positives of the preemptive
tool to the ``search results'' of the exploratory tool, however, with the user input the results can be narrowed down to provide a larger
precision of results given the context. These types of tools stick to previous findings being developers working until something breaks, then
requesting information to fix it.

\begin{description}
	\item[F8\namedlabel{itm:f8}{F8}] Indirect conflict tools should take a curative approach towards indirect conflicts as opposed to preventative
																	 in order to better fit developer activities.
\end{description}

The developers who were surveyed, the following list represents the more common suggestions for how an indirect conflict
curative tools should be implemented.

@@

From the above list, we can see three main areas of focus that researchers of indirect conflicts should spend their time in. The first
is the debugging process through utilizing logs. Showing which recent code changes affect an area of code or how a commit and break a specific
unit test are key methodologies that developers want in order to aid their preferred curative approaches. Developers also want to be able to
understand their software object contracts better or even automatically. Detections of contract changes will play a large part in indirect conflict
and dependency awareness moving forward, especially at the method body and database schema levels. These contract detections also play a role in
the automatic generation of unit tests in order to comply with use case coverage. Finally, and most importantly, developers want tools that
integrate into their current development styles, that being ``work until something breaks''. While a complete evaluation of the curative approach
is beyond the scope of this paper, it is how developers act today and  in the end, tools should be created to aid developers.
